---
title: "DS311 - R Lab Assignment"
author: "Viviana Huang Zheng"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
dim(mtcars)

# Answer:
print("There are total of 11 variables and 32 observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)
lapply(mtcars, unique)

# Answer:
print("There are 6 discrete variables and 5 continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
m = mean(mtcars[, "mpg"])
m
v = var(mtcars[, "mpg"])
v
s = sd(mtcars[, "mpg"])
s


# print(paste("The average of Mile Per Gallon from this data set is ",  , " with variance ",  , " and standard deviation",  , "."))
```
print(paste("The average of Mile Per Gallon from this data set is ", 20.1 , " with variance ", 36.32 , " and standard deviation", 6.03 , "."))


d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
library(ggplot2)
library(dplyr)
mtcars %>% group_by(cyl) %>% summarize(AvgMpg = mean(mpg))

mtcars %>% group_by(gear) %>% summarize(SdMpg = sd(mpg))


```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
mtcars %>% select(cyl, gear) %>% table()

print("The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
plot(PlantGrowth$group, PlantGrowth$weight, main="Plant Growth Per Group", xlab = "group", ylab = "weight", )

```

Result:

=> Report a paragraph to summarize your findings from the plot!
In the control group, the average weight for the plant is around 5.1. The Q1 is ~4.5 whereas the Q3 is 5.4. The minimum weight is 4.2 and the max is 6.2. The first treatment group has an average weight for plants at around 4.5. The Q1 is 4.2 and Q3 is 4.9. The min weight is 3.6 and max weight is 5.9. The second treatment group has an average weight of 5.4. The Q1 is 5.3 and Q3 is 5.8. The min is 4.9 and max is 6.3.  


b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
ggplot(data=mtcars) + geom_histogram(aes(x=mpg), bins=10) + xlab("mpg") + ylab("number of cars") 


print("Most of the cars in this data set are in the class of 15 mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
MA = select(USArrests, Murder, Assault)
pairs(MA, main = 'Murder vs. Assualt')

```

Result:

=> Report a paragraph to summarize your findings from the plot!
Based on the pair plots that display the correlation between Murder and Assault, the two variables are positively correlated. It's not a perfect correlation, r =/1 but the points are all relatively close together, implying there is a correlation between the two. That means that when murder occurs, assault is likely to have also occurred. OR the chance of being murdered increased if you get assaulted. 


***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='C:/Users/vivia/OneDrive/Documents/data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
library(magrittr)
housingData %>% 
     group_by(Neighborhood) %>%   
     summarize(AvgValue = mean(Market.Value.per.SqFt), MedValue = median(Market.Value.per.SqFt)) %>%
     arrange(desc(AvgValue))

```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!


```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!


